---
title: "Intro_Crosstab"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro_Crosstab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Jiner Zheng, Nurah Jaradat, Anna Sophie Peters 
### 21 November 2019

This function calculates and prints a two-way frequency table. Given a data 
frame, a row factor, a column factor, and a type (frequencies, cell percents, 
row percents, or column percents) the function returns a table with labeled rows
and columns, frequencies or percents with the percentage sign depending on 
type, a level labeled \<NA\> if na.rm = FALSE, and a level labeled \<Total\> if 
total = TRUE.
This function coerces x into a factor if it is not already a factor.

```{r setup}
library(qacr)
# crosstab function
crosstab <- function(data, rowvar, colvar, type="freq",
                     total=TRUE, na.rm=FALSE,
                     digits){
  rowvar <- deparse(substitute(rowvar))
  colvar <- deparse(substitute(colvar))
  if (!is.factor(data[[rowvar]]))
    data[rowvar] <- as.factor(data[[rowvar]])
  if (!is.factor(data[[colvar]]))
    data[colvar] <- as.factor(data[[colvar]])
  tb <- table(data[[rowvar]], data[[colvar]])
  names(dimnames(tb)) <- c(rowvar, colvar)
  if (na.rm==TRUE)
    tb <- table(data[[rowvar]], data[[colvar]], useNA = c("ifany"))
  names(dimnames(tb)) <- c(rowvar, colvar)
  if (type=="freq")
    tb
  if (type=="rowpercent")
    tb <- round(prop.table(tb,1), digits+2)
  if (type=="colpercent")
    tb <- round(prop.table(tb,2), digits+2)
  if (type=="cellpercent")
    tb <- round(prop.table(tb),digits+2)
  if (total==TRUE)
    tb <- addmargins(tb, FUN=list("<Total>"=sum), quiet=TRUE)
  result <- list(type=type, digits=digits, tb=tb)
  class(result) <- c("crosstab","list")
  return(result)
}

# print.crosstab function
print.crosstab <- function(x, ...) {
  if(!inherits(x, "crosstab")) stop("Must be a crosstab")
  if(x$type != "freq")
    x$tb <- replace(x$tb, TRUE, sprintf(paste0("%.",x$digits,"f%%"),x$tb*100))
  print.table(x$tb)
}

# Examples of Implementation
crosstab(mtcars, cyl, gear, type = "freq", digits = 2)
crosstab(mtcars, cyl, gear, type = "rowpercent", digits = 3)
```
